A systematic search for transiting planets in the \KT\ data

### 2015-02-15 ms.tex
consistency, todo typography, little crap

The photometry from the \KT\ extension of NASA's \kepler\ mission is
afflicted by systematic effects caused by the relative imprecision of the
telescope pointing and other spacecraft issues.
We present a method for searching these light curves for evidence of
exoplanets by simultaneously fitting for these systematics and the
transit signals of interest.
This method is more computationally expensive than standard search algorithms
but we demonstrate that it can be efficiently implemented and used to robustly
discover transit signals.
We apply this method to the full Campaign 1 dataset and report a list of 38
planet candidates transiting 33 stars.
For all planet candidates, we present posterior distributions on the
properties of each system based strictly on the transit observables.
We also present a list of \dfmtodo{ZZZZ} candidate eclipsing binary systems.

@@

The \paper\ is organized as follows.
In \sect{phot}, we describe our method of extracting aperture photometry from
the calibrated \KT\ postage stamp time series.
In \sect{model} (with details in \app{math}), we describe our data-driven
model for the systematic trends in the photometric light curves and our method
for fitting this model simultaneously with a transit signal.
In \sect{search}, we give the detailed procedure that we use for discovering
and vetting planet candidates.
To quantify the performance and detection efficiency of our pipeline, we test
(in \sect{perform}) the recovery and false alarm rate of synthetic transit
signals, spanning a large range of physical parameters, injected into real
\KT\ light curves.
Finally, in \sect{results}, we present a catalog of 38 planet candidates
orbiting 33 stars---and a supplementary list of \dfmtodo{YYYY} eclipsing binary
candidates---in the publicly available \KT\ Campaign 1 dataset.

@@

The starting point for analysis is the raw pixel data.
We download the full set of 21,703 target pixel files for \KT's Campaign 1
from MAST\footnote{\url{https://archive.stsci.edu/k2/}}.
We extract photometry using fixed circular apertures of varying sizes
centered on the predicted location of the target star based on the world
coordinate system.
\dfmtodo{In detail, the apertures are only approximately circular; each pixel
gets a weight of unity if its center is in the circle, and zero otherwise.}
For each target, we performed the photometry in a set of
circular apertures ranging in radius from 0.5 to
5 pixels (in steps of 0.5 pixels) and---following
\citet{Vanderburg:2014}---chose the aperture size that resulted in the
smallest CDPP \citep{cdpp}
with a 6 hour window.\footnote{Note that although we chose a specific
aperture for each star, photometry for this aperture and every other
aperture size at which we easured is all available online at
\dfmtodo{some URL}.}

@@

The linear search requires hypothesizing a set of transit signals on a
two-dimensional grid in transit time and duration.
For each point in the grid, we use the model described in \sect{model} to
evaluate \emph{the likelihood function} for the transit depth at that time
and duration.
Since the model is linear, the likelihood function for the depth (marginalized
over the model of the systematics) is a Gaussian with analytic amplitude $L$,
mean $\bar{\depth}$, and variance $\delta\bar{\depth}^2$, all derived and given in
\app{math}.
In the linear search, we save these three numbers on a
two-dimensional grid of transit times $\transittime_l$ and durations
$\duration_k$.
The transit time grid spans the full length of Campaign 1 with half hour spacing
and we choose to only test three durations: 1.2, 2.4, and 4.8 hours.
\Fig{linear} shows the mean transit depth $\bar{\depth}$ as a function of
transit time \transittime\ for a light curve with an injected transit signal
with transits at the times indicated by the vertical green bars.

@@

In general, the variable depth model will \emph{always} get a higher
likelihood because it is more flexible.
Therefore, to compare these two models, we use the Bayesian Information
Criterion (BIC).
The traditional definition of the BIC is
\begin{eqnarray}
-\frac{1}{2}\,\BIC_\cdot &=&
    \ln p_\cdot(\{\flux\}_n\,|\,\period,\,\phase,\,\duration)
    - \frac{J}{2} \ln N
\end{eqnarray}
where the likelihood function is maximized, $J$ is an estimate of the model
complexity and $N$ is the effective sample size.
To emphasize that $J$ and $N$ are tuning parameters of the method, we
redefine the BIC as the TIC
\begin{eqnarray}
\TIC_\cdot &=&
    \ln p_\cdot(\{\flux\}_n\,|\,\period,\,\phase,\,\duration) - \alpha
\end{eqnarray}
and we choose $\alpha$ heuristically.
For the K2 Campaign 1 dataset, we find that $\alpha \sim 1250$ leads to reliable
recovery of injected signals while still being fairly insensitive to false
signals.

@@

To measure \emph{centroid offsets}, we start by empirically measuring the
pixel centroid time series for each candidate by modeling the pixels near the
peak as a two-dimensional quadratic and finding the maximum at each time.
This method has been shown to produce higher precision centroid measurements
than center-of-light estimates (Vakili \etal, in preparation).
\dfmtodo{Figure SOMEFIGURE shows the measure $x$ and $y$ pixel coordinate traces for
SOMECANDIDATE.}
Much like the photometry, this signal is dominated by the rigid body motion
of the spacecraft and we can, in fact, model it identically.
Previously (\eqalt{linear-model}), we modeled the light curve as a linear
combination of ELCs and a simple box transit model at a given period, phase,
and duration.
Under this model, the maximum likelihood depth can be computed analytically.
If we apply \emph{exactly the same model} to the centroid trace, the ``depth''
that we compute becomes the centroid motion in transit in units of pixels.
Since the motions won't necessarily point in a consistent direction across
transits, we treat each transit independently and report the average offset
amplitude weighted by the precision of each measurement.
To compute the significance of a centroid offset, we bootstrap the offset
amplitude for models at the same period and duration but randomly oriented
phases.
If the centroid measured for the candidate transit is substantially larger
than the random realizations, we label the candidate as a false positive.
In practice, the precision of the centroid measurements isn't sufficient to
robustly reject many candidates but two candidates---EPIC 201202105 and EPIC
201632708---have offsets large enough that we remove them from our final
catalog.
Figure SOMEFIGURE shows the distribution of centroid offsets across the full
catalog.

@@

{\item Draw the number of transiting planets based on the observed
multiplicity distribution of KOIs (CITE).}

@@

Out of the 21,703 Campaign 1 light curves, our pipeline returns 741 signals
that pass the signal-to-noise and period cuts.
After hand vetting by the two first authors, this list is reduced to 101
convincing astrophysical transit candidates.
Of these, 38 signals---in 33 light curves---have no visible secondary
eclipse and these planet candidates are listed in \tab{cand}.
Two were previously published \citep{Crossfield:2015} but, to our knowledge,
the remaining list is currently unpublished.
The third planet published by \citet{Crossfield:2015} is found as the third
signal for that target by our pipeline but it falls just below our
signal-to-noise cut so we left it out of the catalog for consistency.
This suggests that a less conservative cut in signal-to-noise and more
aggressive machine vetting could yield a much more complete catalog at smaller
radii and longer periods even with the existing dataset.

@@

SOMEFIGURE shows the observed distribution of planet candidates in our
catalog.
Using this and the detection efficiency function computed in the previous
section, we compute the occurrence rate as a function of period by ignoring
the measurement uncertainties and following the method described in the
Appendix of \citet{dfm}.
This function is shown in SOMEFIGURE and IS THIS CONSISTENT INTERESTING ETC?

In \papertwo, we will characterize the stars for each of our candidates in
detail but for now it's worth noting that about half (IS THIS TRUE) of our
planet candidates are orbiting M-type stars.
If this rate remains robust after stellar characterization and if these
numbers are representative of the \KT\ yields in upcoming Campaigns, this
will substantially increase the number of planets known to transit cool stars!

@@

The flexible systematics model we employ is a 150-parameter linear combination of PCA components
derived from the full set of 20,000 stellar lightcurves.
That is, it presumes that the systematics afflicting each star are shared in
some way across other stars.
It is our belief---although not a strict assumption of our model---that these
systematics are caused primarily by pointing drifts, or movements of the
pixels in the focal plane relative to the stars.
In principle, if the systematics \emph{are} dominated by pointing issues, the
systematics model could require only three parameters---three Euler
angles---not 150 amplitudes.
However, because (as the pointing drifts) each star sees it's own unique
local patch of flat-field variations, the mapping from pointing drifts to
brightness variations can be extremely non-linear.
Furthermore, because when the pointing is moving fast there is effectively
a smearing of the point-spread function, there are effects keyed to the
time derivative of the Euler angles as well.
The large number (150) of linear coefficients gives the linear model the
freedom to model complex non-linear behavior; we are trading off parsimony
in parameters with the enormous computational advantages of maintaining
linearity (and therefore also convexity).
The computational advantages of the linear model are three-fold:
Convexity obviates searching in parameter space for alternative modes;
linear least-squares optimization can be performed with simple linear algebra;
given Gaussian uncertainties and uninformative priors, marginalizations over
the linear parameters also reduces to pure linear algebra.

@@

\begin{figure}[p]
\begin{center}
\includegraphics{figures/pca.pdf}
\end{center}
\caption{%
The top 10 eigen light curves (ELCs) generated by running principal component
analysis on all the aperture photometry from Campaign 1.
\figlabel{pca}}
\end{figure}

>>>>

The photometry from the \KT\ extension of NASA's \kepler\ mission is
afflicted by systematic effects caused by the relative imprecision of the
telescope pointing and other spacecraft issues.
We present a method for searching these light curves for evidence of
exoplanets by simultaneously fitting for these systematics and the
transit signals of interest.
This method is more computationally expensive than standard search algorithms
but we demonstrate that it can be efficiently implemented and used to robustly
discover transit signals.
We apply this method to the full Campaign~1 dataset and report a list of 38
planet candidates transiting 33 stars.
For all planet candidates, we present posterior distributions on the
properties of each system based strictly on the transit observables.
We also present a list of \dfmtodo{ZZZZ} candidate eclipsing binary systems.

@@

The \paper\ is organized as follows.
In \sect{phot}, we describe our method of extracting aperture photometry from
the calibrated \KT\ postage stamp time series.
In \sect{model} (with details in \app{math}), we describe our data-driven
model for the systematic trends in the photometric light curves and our method
for fitting this model simultaneously with a transit signal.
In \sect{search}, we give the detailed procedure that we use for discovering
and vetting planet candidates.
To quantify the performance and detection efficiency of our pipeline, we test
(in \sect{perform}) the recovery and false alarm rate of synthetic transit
signals, spanning a large range of physical parameters, injected into real
\KT\ light curves.
Finally, in \sect{results}, we present a catalog of 38 planet candidates
orbiting 33 stars---and a supplementary list of \dfmtodo{YYYY} eclipsing binary
candidates---in the publicly available \KT\ Campaign~1 dataset.

@@

The starting point for analysis is the raw pixel data.
We download the full set of 21,703 target pixel files for \KT's Campaign~1
from MAST\footnote{\url{https://archive.stsci.edu/k2/}}.
We extract photometry using fixed circular apertures of varying sizes
centered on the predicted location of the target star based on the world
coordinate system.
\dfmtodo{In detail, the apertures are only approximately circular; each pixel
gets a weight of unity if its center is in the circle, and zero otherwise.}
For each target, we performed the photometry in a set of
circular apertures ranging in radius from 0.5 to
5 pixels (in steps of 0.5 pixels) and---following
\citet{Vanderburg:2014}---chose the aperture size that resulted in the
smallest CDPP \citep{cdpp}
with a 6 hour window.\footnote{Note that although we chose a specific
aperture for each star, photometry for this aperture and every other
aperture size at which we easured is all available online at
\dfmtodo{some URL}.}

@@

The linear search requires hypothesizing a set of transit signals on a
two-dimensional grid in transit time and duration.
For each point in the grid, we use the model described in \sect{model} to
evaluate \emph{the likelihood function} for the transit depth at that time
and duration.
Since the model is linear, the likelihood function for the depth (marginalized
over the model of the systematics) is a Gaussian with analytic amplitude $L$,
mean $\bar{\depth}$, and variance $\delta\bar{\depth}^2$, all derived and given in
\app{math}.
In the linear search, we save these three numbers on a
two-dimensional grid of transit times $\transittime_l$ and durations
$\duration_k$.
The transit time grid spans the full length of Campaign~1 with half hour spacing
and we choose to only test three durations: 1.2, 2.4, and 4.8 hours.
\Fig{linear} shows the mean transit depth $\bar{\depth}$ as a function of
transit time \transittime\ for a light curve with an injected transit signal
with transits at the times indicated by the vertical green bars.

@@

In general, the variable depth model will \emph{always} get a higher
likelihood because it is more flexible.
Therefore, to compare these two models, we use the Bayesian Information
Criterion (BIC).
The traditional definition of the BIC is
\begin{eqnarray}
-\frac{1}{2}\,\BIC_\cdot &=&
    \ln p_\cdot(\{\flux\}_n\,|\,\period,\,\phase,\,\duration)
    - \frac{J}{2} \ln N
\end{eqnarray}
where the likelihood function is maximized, $J$ is an estimate of the model
complexity and $N$ is the effective sample size.
To emphasize that $J$ and $N$ are tuning parameters of the method, we
redefine the BIC as the TIC
\begin{eqnarray}
\TIC_\cdot &=&
    \ln p_\cdot(\{\flux\}_n\,|\,\period,\,\phase,\,\duration) - \alpha
\end{eqnarray}
and we choose $\alpha$ heuristically.
For the K2 Campaign~1 dataset, we find that $\alpha \sim 1250$ leads to reliable
recovery of injected signals while still being fairly insensitive to false
signals.

@@

To measure \emph{centroid offsets}, we start by empirically measuring the
pixel centroid time series for each candidate by modeling the pixels near the
peak as a two-dimensional quadratic and finding the maximum at each time.
This method has been shown to produce higher precision centroid measurements
than center-of-light estimates (Vakili \etal, in preparation).
\dfmtodo{Figure SOMEFIGURE shows the measure $x$ and $y$ pixel coordinate traces for
SOMECANDIDATE.}
Much like the photometry, this signal is dominated by the rigid body motion
of the spacecraft and we can, in fact, model it identically.
Previously (\eqalt{linear-model}), we modeled the light curve as a linear
combination of ELCs and a simple box transit model at a given period, phase,
and duration.
Under this model, the maximum likelihood depth can be computed analytically.
If we apply \emph{exactly the same model} to the centroid trace, the ``depth''
that we compute becomes the centroid motion in transit in units of pixels.
Since the motions won't necessarily point in a consistent direction across
transits, we treat each transit independently and report the average offset
amplitude weighted by the precision of each measurement.
To compute the significance of a centroid offset, we bootstrap the offset
amplitude for models at the same period and duration but randomly oriented
phases.
If the centroid measured for the candidate transit is substantially larger
than the random realizations, we label the candidate as a false positive.
In practice, the precision of the centroid measurements isn't sufficient to
robustly reject many candidates but two candidates---EPIC 201202105 and EPIC
201632708---have offsets large enough that we remove them from our final
catalog.
Figure~\dfmtodo{SOMEFIGURE} shows the distribution of centroid offsets across the full
catalog.

@@

{\item Draw the number of transiting planets based on the observed
multiplicity distribution of KOIs (\dfmtodo{CITE}).}

@@

Out of the 21,703 Campaign~1 light curves, our pipeline returns 741 signals
that pass the signal-to-noise and period cuts.
After hand vetting by the two first authors, this list is reduced to 101
convincing astrophysical transit candidates.
Of these, 38 signals---in 33 light curves---have no visible secondary
eclipse and these planet candidates are listed in \tab{cand}.
Two were previously published \citep{Crossfield:2015} but, to our knowledge,
the remaining list is currently unpublished.
The third planet published by \citet{Crossfield:2015} is found as the third
signal for that target by our pipeline but it falls just below our
signal-to-noise cut so we left it out of the catalog for consistency.
This suggests that a less conservative cut in signal-to-noise and more
aggressive machine vetting could yield a much more complete catalog at smaller
radii and longer periods even with the existing dataset.

@@

Figure~\dfmtodo{SOMEFIGURE} shows the observed distribution of planet candidates in our
catalog.
Using this and the detection efficiency function computed in the previous
section, we compute the occurrence rate as a function of period by ignoring
the measurement uncertainties and following the method described in the
Appendix of \citet{dfm}.
This function is shown in Figure \dfmtodo{SOMEFIGURE and IS THIS CONSISTENT INTERESTING ETC?}

In \papertwo, we will characterize the stars for each of our candidates in
detail but for now it's worth noting that about half \dfmtodo{(IS THIS TRUE?)} of our
planet candidates are orbiting M-type stars.
If this rate remains robust after stellar characterization and if these
numbers are representative of the yields in upcoming \KT\ Campaigns, this
will substantially increase the number of planets known to transit cool stars!

@@

The flexible systematics model we employ is a 150-parameter linear combination of PCA components
derived from the full set of 21,703 stellar lightcurves.
That is, it presumes that the systematics afflicting each star are shared in
some way across other stars.
It is our belief---although not a strict assumption of our model---that these
systematics are caused primarily by pointing drifts, or movements of the
pixels in the focal plane relative to the stars.
In principle, if the systematics \emph{are} dominated by pointing issues, the
systematics model could require only three parameters---three Euler
angles---not 150 amplitudes.
However, because (as the pointing drifts) each star sees it's own unique
local patch of flat-field variations, the mapping from pointing drifts to
brightness variations can be extremely non-linear.
Furthermore, because when the pointing is moving fast there is effectively
a smearing of the point-spread function, there are effects keyed to the
time derivative of the Euler angles as well.
The large number (150) of linear coefficients gives the linear model the
freedom to model complex non-linear behavior; we are trading off parsimony
in parameters with the enormous computational advantages of maintaining
linearity (and therefore also convexity).
The computational advantages of the linear model are three-fold:
Convexity obviates searching in parameter space for alternative modes;
linear least-squares optimization can be performed with simple linear algebra;
given Gaussian uncertainties and uninformative priors, marginalizations over
the linear parameters also reduces to pure linear algebra.

@@

\begin{figure}[p]
\begin{center}
\includegraphics{figures/pca.pdf}
\end{center}
\caption{%
The top 10 eigen light curves (ELCs) generated by running principal component
analysis on all the aperture photometry from Campaign~1.
\figlabel{pca}}
\end{figure}
